\documentclass{tufte-handout}

\usepackage{xcolor}

% set image attributes:
\usepackage{graphicx}
\graphicspath{ {images/} }

% set hyperlink attributes
\hypersetup{colorlinks}

% create environment for bottom paragraph:
\newenvironment{bottompar}{\par\vspace*{\fill}}{\clearpage}

\usepackage{enumerate}

% set table attributes
\usepackage{tabu}
\usepackage{booktabs}

% ============================================================

% define the title
\title{SOC 4015/5050: Lab-13 - Regression Diagnostics}
\author{Christopher Prener, Ph.D.}
\date{Fall 2018}

% ============================================================

\begin{document}

% ============================================================

\maketitle % generates the title

% ============================================================

\vspace{5mm}
\section{Directions}
Please complete all steps below. All work should be uploaded to your GitHub assignment repository by 4:15pm on Monday, December 3\textsuperscript{rd}, 2018. All data can be obtained from the \texttt{testDriveR} package's \texttt{auto17} data set.

\vspace{5mm}
\section{Analysis Development}
Using RStudio and your operating system's file manager, create an R Project in the \textit{existing} directory in your assignments repository named \texttt{Lab-13}. Add a \texttt{README.md} file, notebook, and all necessary folders before beginning.\sidenote{This initial section follows the project workflow that is available in the \texttt{lecture-03} repo!}

\vspace{3mm}
\section{Fit An Initial Model}
\begin{enumerate}
\item Execute a multivariate regression model that shows how \texttt{displ} affects \texttt{fuelCost} controlling for characteristics of the engine (\texttt{cyl} and \texttt{gears}) and highway fuel efficiency (\texttt{hwy}).
\end{enumerate}

\vspace{3mm}
\section{Assess Model}
\begin{enumerate}
\setcounter{enumi}{1}
\item Check model for non-linearity in the relationship between \textit{x} variables and \textit{y}.
\item Check the residuals for normality.
\item Check the residuals for homoskedastic errors.
\item Check the residuals for auto-correlation.
\item Check the independent variables for multi-collinearity.
\item Summarize your findings - do you have concerns about how the model is specified? Should variables be removed or added?
\end{enumerate}

\vspace{3mm}
\section{Check for Unusual Observations}
\begin{enumerate}
\setcounter{enumi}{7}
\item Check for outliers.
\item Check for observations with high leverage values.
\item Check for observations with high influence values.
\item Summarize your findings - do you have concerns about unusual observations? Should observations be removed? If so, which ones?
\end{enumerate}

\vspace{3mm}
\section{Re-Fit Model}
\begin{enumerate}
\setcounter{enumi}{11}
\item Re-fit the model based on your findings from the previous two sections. 
\item Compare your initial model and the newly re-fit model. How do the betas differ (if at all)? 
\end{enumerate}

% ============================================================
\end{document}